---
title: "Homework 3"
author: "Hongyu Yu"
output: pdf_document
---

# BIOS 611 Homework 3

### Question 1

```{r}
require(tidyverse)

# read in data and print problem
raw.data <- read_csv("NYC_Dog_Licensing_Dataset_20240918.csv", show_col_types = F)
pr <- problems(raw.data)
print(pr)

# rename the columns
data <- raw.data %>% 
  rename(
  name=AnimalName,
  gender=AnimalGender,
  birth_year=AnimalBirthYear, 
  breed=BreedName, 
  zipcode=ZipCode, 
  issue_date=LicenseIssuedDate, 
  expiration_date=LicenseExpiredDate,
  extract_year=`Extract Year`
)

# reduce to complete case
cat(sprintf("%d rows before filtering complete cases", nrow(data)))
data <- data %>% filter(complete.cases(.))
cat(sprintf("%d rows after filtering complete cases", nrow(data)))
```

### Question 2

By sorting the duplicated rows by their name and the count of their occurrence, we found that not all duplicated rows corresponds to dog names. Instead, some rows have dog names denoted by a ".", and other rows have names "UNKNOWN", "NAME NOT PROVIDED", "NAME", and "NONE". Drop those rows and save the cleaned data.

```{r}
distinct.data <- data %>% distinct()
write_csv(distinct.data, "distinct_data.csv")
cat(sprintf("%d distinct rows", nrow(distinct.data)))
# extract duplicated rows
duplicated <- data %>% group_by_all() %>% filter(n() > 1) %>% distinct()
print(duplicated)
name.count <- duplicated %>%
  group_by(name) %>%
  summarise(n=n())
print(name.count[1:20,])
print(arrange(name.count, desc(n))[1:20,])

deduplicated <- distinct.data %>%
  filter(! name %in% c(".", "UNKNOWN", "NAME NOT PROVIDED", "NAME", "NONE")) %>%
  write_csv("deduplicated.csv")
```

### Question 3

For the second plot, only years with at least 20 dogs are plotted.

```{r}
ggplot(deduplicated, aes(birth_year)) + geom_histogram(stat="count") +
  labs(x="Birth year", y="Dog count", title="All year dog count")
ggsave("dog_count_all.pdf")

nontrivial.years <- deduplicated %>% 
  group_by(birth_year) %>%
  summarise(count=n()) %>%
  filter(count > 20)
deduplicated %>% 
  filter(birth_year %in% nontrivial.years$birth_year) %>%
  ggplot(aes(birth_year)) + geom_histogram(stat="count") + 
  labs(x="Birth year", y="Dog count", title="Nontrivial year dog count")
ggsave("dog_count_nontrivial.pdf")
```

### Question 4

```{r}
sprintf("%d rows before filtered by name, gender, birth year, breed, and zip code",
        nrow(deduplicated))
unique.data <- deduplicated %>%
  distinct(name, gender, birth_year, breed, zipcode)
sprintf("%d rows after filtered by name, gender, birth year, breed, and zip code",
        nrow(unique.data))
write_csv(unique.data, "unique.csv")
```

### Question 5

There are two dog names with the same rate for 1999, which are BUDDY and MAX. The most popular dog name in 2023 is LUNA after removing ., NONE, NAME, NAME NOT PROVIDED, and UNKNOWN.

```{r}
dogs.by.year <- unique.data %>%
  group_by(birth_year) %>%
  summarise(count=n())

dogs.by.year.name <- unique.data %>%
  group_by(birth_year, name) %>%
  summarise(name_count=n(), .groups = "drop")

joined.count <- dogs.by.year %>% 
  right_join(dogs.by.year.name, by=c("birth_year")) %>%
  mutate(rate=name_count/count)

df.1999 <- joined.count[joined.count$birth_year==1999,]
max.rate.1999 <- df.1999$name[df.1999$rate==max(df.1999$rate)]
cat("The most popular dog names in 1999 are/is:", 
    paste(max.rate.1999, collapse = ", "))

df.2023 <- joined.count[joined.count$birth_year==2023,]
max.rate.2023 <- df.2023$name[df.2023$rate==max(df.2023$rate)]
cat("The most popular dog names in 2023 are/is:", 
    paste(max.rate.2023, collapse = ", "))
```

### Question 6

```{r}
joined.count %>% 
  filter(name %in% c(max.rate.1999, max.rate.2023)) %>%
  ggplot(aes(x=birth_year, y=rate, color=name)) +
  geom_line() + 
  labs(x="Birth year", y="Rate")
ggsave("max_name.pdf")
```

### Makefile

```{make}
build: dog_count_all.pdf dog_count_nontrivial.pdf \
	max_name.pdf

clean:
	rm *.pdf
	rm distinct_data.csv
	rm unique.csv
	rm deduplicated.csv

distinct_data.csv: preprocess.r
	Rscript preprocess.r

unique.csv: distinct_data.csv unique.r
	Rscript unique.r

dog_count_all.pdf dog_count_nontrivial.pdf: distinct_data.csv plot1.r
	Rscript plot1.r

max_name.pdf: unique.csv plot2.r
	Rscript plot2.r
```

